---
title: "Assignment 3"
author: "Ashish Mhatre"
date: "04/02/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem Statement

Perform data analysis and visualization on Wine dataset, Farmer Market dataset and Airline delays dataset using data wrangling techniques.

## Import Libraries

Importing all required modules 

```{r libraries, include=FALSE}
library(tidyr) #Importing all dependencies
library(dplyr)
library(lubridate)
library(magrittr)
library(ggplot2)
library(gridExtra)
library(stringr)

```

## Task 1

Perform 5 subtask on Wine dataset.

### Subtask 1

Find region in france having the highest average price.

```{r subtask1, echo=FALSE, warning=FALSE}
wine = read.csv("C:/Users/Ashish Mhatre/Desktop/CVA/Assigments/Assignment 3/wine_data.csv")

wine%>%
  filter(country == 'France')%>% 
  select(region_1, price)%>%
  group_by(region_1)%>% #Group by region
  summarise_at(vars(price),list(Average_Price = mean))%>% #Take Average of prices
  drop_na(Average_Price)%>%
  arrange(desc(Average_Price))%>%
  slice(1) #Taking the top region

              

```

Montrachet region has the highest average price of 601.1818

### subtask 2

Find average price by designation

```{r subtask2, echo=FALSE, warning=FALSE}
wine%>%
  select(designation, price)%>%
  group_by(designation)%>% #Group by designation
  summarise_at(vars(price),list(Average_Price = mean))%>% #Take average of prices 
  drop_na(Average_Price)%>%
  arrange(desc(Average_Price))
  


```

Following data set view shows the average price by designation

### Subtask 3

Find the variety having the highest average price.

```{r subtask3, echo=FALSE, warning=FALSE}

wine%>%
  select(variety, price)%>%
  group_by(variety)%>% #Group by variety
  summarise_at(vars(price),list(Average_Price = mean))%>% #Take average of prices
  drop_na(Average_Price)%>%
  arrange(desc(Average_Price))%>%
  slice(1)
  


```

Carbernet-Shiraz has the highest price of 150

## Subtask 4

Display top 7 variety by count frequency.

```{r subtask4, echo=FALSE, warning=FALSE}


wine%>%
  select(variety, price)%>%
  group_by(variety)%>%  #Group by variety
  summarise( Count = n())%>%  #Count the frequency
  arrange(desc(Count))%>%
  slice(1:7) #display top 7



```

Following dataset shows the top 7 variety by count frequency.

### Subtask 5 

Finding the number of wines which are 20 years old


```{r subtask5, echo=FALSE, warning=FALSE}

win_20 <- wine%>%
  select(designation)%>%
  filter(str_detect( designation,regex('((20)+(\\s|-)*((Y|years?)|(Y|yr.)))|(20th)') ) == TRUE ) #Set regex pattern to match the 20 year string parts

cat("Count of 20 years old Wine is ", nrow(win_20)) #display the number of rows by nrow

```

There are 84 wines which are 20 year old.

## Task 2

Generate wide table showing number of farmers market by state and month of year.

```{r Task 2, echo=FALSE, warning=FALSE}

fm = read.csv("C:/Users/Ashish Mhatre/Desktop/CVA/Assigments/Assignment 3/fm.csv")

fm$Season1Date <- as.Date(fm$Season1Date,'%m/%d/%y' ) #Convert to Datetime

fm%>%
  select(State,Season1Date)%>%
  drop_na(Season1Date)%>% #Drop NA values
  group_by(State, Months = months(Season1Date))%>% #Group by state first then months
  summarise( count = n(), .groups='drop')%>% #count the frequency
  arrange( factor( Months, levels = c('January', 'February', 'March', 'April', 'May', 'June', 'July', 'August', 'September', 'October', 'November', 'December')))%>% #Set the order of categorical variable
  drop_na(count)%>%
  pivot_wider(names_from = Months, values_from = count)%>% #Pivot wider on Months column
  replace_na(list(January = 0, February = 0, March = 0, April = 0, May = 0,  June = 0, July = 0, August = 0, September = 0, October = 0, November = 0, December = 0)) #Replace NA with 0 in Months columns

```

We can see the number of farmers market for each month for a particular state in a single row.

## Task 3 

Showing the number of active farmer markets depending on the Updatetime column by month for each city in state of california

```{r task 3, echo=FALSE, warning=FALSE}

 fm$updateTime <- as.Date(fm$updateTime, '%m/%d/%Y' )

fm%>%
  filter(State == 'California')%>%
  select(city, updateTime)%>%
  drop_na(updateTime)%>%
  group_by(city, Update_month = month(updateTime))%>% #Group by city and Update_Months
  summarise(Active_Users = n(), .groups='drop')%>%
  arrange(desc(Active_Users))



```

Following dataset view provides us with the number of active farmers market for each city of california for each month.

## Task 4 

Plot a bar chart to display the total delay in mins by each carrier

```{r task 4, echo=FALSE, warning=FALSE}

delay = read.csv("C:/Users/Ashish Mhatre/Desktop/CVA/Assigments/Assignment 3/airlines_delay.csv")

delay$year <- as.character(delay$year) # to convert in Categorical data

total_delay <- delay%>%
               select(carrier,arr_delay)%>%
               group_by(carrier)%>% #Group by carrier
               summarise( total_delay = sum(arr_delay, na.rm = TRUE) ) %>%  #Add all the delays 
               arrange(desc(total_delay))
               

ggplot(data=total_delay, aes(x=reorder(carrier,total_delay), y=total_delay)) + 
  geom_bar(stat="identity", alpha = 0.6) + ggtitle("Fig1. Carrier vs Total Delay") + xlab('Carriers') + ylab('Total delay in mins')
   

```

We can see the distribution of total delays for each carrier where, x-axis represents the carrier name and y-axis represents the total delay in mins.

## Task 4.1

Plot a stacked plot showing the total delay for each year stacked for a particular airline carrier.

```{r task4.1, echo=FALSE, warning=FALSE}

delay_year <- delay%>%
               select(year,carrier,arr_delay)%>%
               group_by(carrier,year)%>% #Group by carrier first then year
               summarise( total_delay = sum(arr_delay, na.rm = TRUE), .groups='drop' )%>% # Count all the sum by year and carrier
               arrange(desc(total_delay))

join_df <- inner_join(total_delay,delay_year,by = 'carrier') #Perfrom inner join

ggplot(data = join_df, aes(x=reorder(carrier,-total_delay.x), y=total_delay.y, fill= as.factor(year))) +
      geom_bar(stat="identity") + 
      scale_fill_brewer(palette = 'Paired')
 
```

The following graph shows carrier vs total delay by year, where x-axis represents the airline carrier and y-axis represent the time delay in mins. each year delay is represented with an different color as a stacked bar.

# Conclusion

1. Performed following subtask :-

a. Montrachet region has the highest average price of 601.1818.

b. Displayed the average price by designation where in designation Clos du Mesnil has the highest average price 1400.00000.	

c. Carbernet-Shiraz variety has the highest price of 150.

d. using group_by and summaries clause displayed data set view for the top 7 variety by count frequency.

e. using str_detect and reg-ex we found out that There are 84 wines which are 20 year old.

2. Generated view to see the number of farmers market for each month for a particular state in a single row using pivot_wider.

3. computed Number of active farmers market for each city of California for each month.where we found that San Francisco
in the month of July had the most number of active markets that is 8.

4.We can see the distribution of total delays for each carrier where, x-axis represents the carrier name and y-axis represents the total delay in mins. where southwest airlines co. has the highest total delay compared to all airlines, 
also using (fill=year) argument in ggplot we created a stacked bar graph adding details for total delay by each year.


